{ \fontfamily{ptm}\selectfont Nesterov's First Order Method Accelerations: Unifications, Applications and Numerial Experiments }

### 2025-02-07 Notebook LaTEX/Proposal/presentation.tex
>w<, changes to proposal.

% SLIDE INFORMATION ============================================================
\author{Hongda Li}
\title[Thesis Proposal Talk]{First Order Nonsmooth Optimization: Catalyst Acceleration and Unifying Nesterov's Acceleration}
% \newcommand{\email}{lalala@lala.la}
\institute[UBCO]{
    University of British Columbia Okanagan
}
\date{\today}
\subject{Nesterov's acceleration and its applications}

@@

@@

                \item It's an early attempt at putting accelerated inexact proximal point method into a practical setting. 

@@

\section{Content of the draft paper}

@@

                \item We proved the equivalences in our draft papers and the convergence claim from previous applies to all the equivalent forms of R-WAPG which will follow. 

@@

@@

        \begin{frame}{Catalyst meta acceleration}

>>>>

% SLIDE INFORMATION ============================================================
\author{Hongda Li}
\title[Thesis Proposal Talk]{Nesterov's First Order Method Accelerations: Unifications, Applications  and Numerial Experiments}
% \newcommand{\email}{lalala@lala.la}
\institute[UBCO]{
    University of British Columbia Okanagan
}
\date{\today}
\subject{Nesterov's acceleration and its applications}

@@

    \begin{frame}{Introduction}
        \begin{block}{Upcoming Contents: }
            \begin{itemize}
                \item Notations, proximal gradient inequality, gradient mapping. 
                \item Nesterov's estimating sequence.
                \item Introducing R-WAPG (Content of our draft paper). 
                \item Introducing the Catalyst metal acceleration framework. 
            \end{itemize}

@@

                \item It puts accelerated inexact proximal point method into a practical setting. 

@@

\section{Content of the draft paper}
    \begin{frame}{Contents of our draft papers}
        \begin{block}{Upcoming contents}
            We explain results from our relaxed, weak accelerated proximal gradient (R-WAPG). 
            \begin{enumerate}
                \item The R-WAPG sequence and the method of R-WAPG. 
                \item The generic convergence of R-WAPG. 
                \item Three equivalent forms of R-WAPG. 
                \item Specialized convergence results with for specific instances of FISTA variants in the literatures. 
                \item A parameter free formulation of R-WAPG and the results of our preliminary numerical experiments. 
                \item Nesterov's idea of strong convexity transfer and the R-WAPG framework. 
            \end{enumerate}
        \end{block}

@@

                \item We proved the upcoming equivalences forms of R-WAPG in the draft papers. 
                \item Previous claimed generic convergence results will apply to all equivalent forms of R-WAPG will now come. 

@@

    \begin{frame}{Selected contents from Catalyst Meta Accelerations}
        \begin{block}{Upcoming Contents}
                \item Notations, assumptions. 
                \item Inexact proximal point evaluations criteria. 
                \item Structure of Catalyst accelerations and inner/out loop complexity results from Lin et al's first Catalyst paper \cite{lin_universal_2015}. 
                \item Catalyst acceleration with relative error crition, Lin et al's second paper \cite{lin_catalyst_2018}.
                \item Comments on future works for Catalyst by Necoara et al. \cite{necoara_linear_2019} and our questions about the Catalyst Acceleration Framework. 

@@

        \begin{frame}{Catalyst Meta Acceleration}

### 2025-02-07 Notebook LaTEX/Proposal/proposal.tex
>w<, changes to proposal.

                \item $M, N$ are constants. They define matrix $A \in \RR^{M\times N}$ which has entires of i.i.d random variable taken from a standard normal distribution. 
                \item $L, \mu$, are Lipschitz constant and the strong convexity constant for the smooth part of the objective which are not known prior. Hence they are estimated by $A$ by $\mu = 1/\Vert (A^TA)^{-1}\Vert$ and $L = \Vert A^TA\Vert$ prior to experiment to assist the V-FISTA algorithm. 

@@

\section{Catalyst accelerations}\label{sec:catalyst}
    We introduce assumptions and notations for the Catalyst Meta Acceleration Frameworks. 
    \begin{assumption}\label{ass:catalyst1}
        Given any $\beta > 0$ and $y \in \RR^n$, and $F: \RR^n\rightarrow \overline \RR$ is $\mu \ge 0$ strongly convex and closed. 
        Assume that minimizer exists for $F$ and the minimum is $F^*$. 
        For all $x,y\in \RR^n, \beta > 0$ define the model function: 
        \begin{align*}
            \mathcal M^{\beta^{-1}}_F(x; y) &:= 
            F(x) + \frac{\beta}{2}\Vert x - y\Vert^2.
        \end{align*}
        We define the Moreau Envelope at $y \in \RR^n$ to be $\mathcal M^*_{F,\beta^{-1}}(y) := \min_{x\in \RR^n} \mathcal M_F^{\beta^{-1}}(x; y)$ and the resolvent operator for subgradient of $F$ to be $\mathcal J_{\beta^{-1}F}$, i.e: which is also called the proximal operator. 
        \par
        We invent this notations because there are no consistent notations in the literatures and several of the Catalyst papers all use dratically different notations.  
    \end{assumption}
    \begin{definition}[Absolute termination criterion C1]\label{def:catalyst-termination-c1}
        Take $F$ as given by Assumption \ref{ass:catalyst1}.
        For $\epsilon > 0, \kappa > 0$ and $x \in \RR^n$, the absolute criterion C1 characterizes the set of inexact proximal iterates by: 
        \begin{align*}
            \mathcal J_{\kappa^{-1} F}^\epsilon (x) := 
            \left\lbrace
                y \in \RR^n \left | \; 
                        \mathcal M^{1/\kappa}_F (y; x) - 
                        \mathcal M^*_{F, 1/\kappa}(x) \le \epsilon
                \right.
            \right\rbrace. 
        \end{align*}
    \end{definition}
    \begin{remark}
        Setting $\epsilon = 0$, we have the exact definition of the exact resolvent given as $\mathcal J_{\beta^{-1}}y = \mathcal J^0_{\beta^{-1}}y$. 
    \end{remark}
    \textbf{The remainder of this section is organized as the follow}. 
    Section \ref{ssec:catalyst-intro} gives the main ideas behind the Catalyst Acceleration Framework. 
    It follows Lin et al.'s first paper \cite{lin_universal_2015} on Catalyst Acceleration and summarizes the convergence analysis of the algorithm. 
    Section \ref{ssec:catalyst-second-paper} introduces Lin et al.'s second paper \cite{lin_catalyst_2018} on Catalyst which considers the improved relative error condition stated in Definition \ref{def:catalyst-termination-c2}. 
    The relative error condition improves the total complexity and inspired nonconvex gradient based 4WD Catalyst by Paquette et al. \cite{paquette_catalyst_2018}.

@@

        Inspired by accelerated proximal point Method (APPM) from Güler \cite{guler_new_1992}, and Inexact Proximal Point Method (IPPM) of Rockafellar 1976 \cite{rockafellar_monotone_1976}, Lin et al. \cite{lin_universal_2015} proposed a generic method taking inspirations from the convergence claims of APPM to accelerate Variance Reduced Method (VRM). 

@@

            \begin{theorem}[Outer loop convergence convex but not strongly convex]\label{thm:erro-seq-outer-cnvx}\;\\

@@

                Assume that $F$ is convex but with strong convexity constant $\mu = 0$. 

@@

                VRMs listed earlier statsifies this assumption. 

@@

            \begin{proposition}[Inner loop, context but not strongly convex]\label{prop:inner-loop-complexity-cnvx}

>>>>

                \item $M, N$ are constants. They define matrix $A \in \RR^{M\times N}$ which has entries of i.i.d random variable taken from a standard normal distribution. 
                \item $L, \mu$, are Lipschitz constant and the strong convexity constant for the smooth part of the objective which are not known prior. Hence, they are estimated by $A$ by $\mu = 1/\Vert (A^TA)^{-1}\Vert$ and $L = \Vert A^TA\Vert$ prior to experiment to assist the V-FISTA algorithm. 

@@

\section{Catalyst accelerations}\label{sec:catalyst}
    We introduce assumptions and notations for the Catalyst Meta Acceleration Frameworks. 
    \begin{assumption}\label{ass:catalyst1}
        Given any $\beta > 0$ and $y \in \RR^n$, and $F: \RR^n\rightarrow \overline \RR$ is $\mu \ge 0$ strongly convex and closed. 
        Assume that minimizer exists for $F$ and the minimum is $F^*$. 
        For all $x,y\in \RR^n, \beta > 0$ define the model function: 
        \begin{align*}
            \mathcal M^{\beta^{-1}}_F(x; y) &:= 
            F(x) + \frac{\beta}{2}\Vert x - y\Vert^2.
        \end{align*}
        We define the Moreau Envelope at $y \in \RR^n$ to be $\mathcal M^*_{F,\beta^{-1}}(y) := \min_{x\in \RR^n} \mathcal M_F^{\beta^{-1}}(x; y)$ and the resolvent operator for subgradient of $F$ to be $\mathcal J_{\beta^{-1}F}$, i.e: which is also called the proximal operator. 
        \par
        We invent the notations because there are no consistent notations in the literatures and several of the Catalyst papers all use different notations.  
    \end{assumption}
    \begin{definition}[Absolute termination criterion C1]\label{def:catalyst-termination-c1}
        Take $F$ as given by Assumption \ref{ass:catalyst1}.
        For $\epsilon > 0, \kappa > 0$ and $x \in \RR^n$, the absolute criterion C1 characterizes the set of inexact proximal iterates by: 
        \begin{align*}
            \mathcal J_{\kappa^{-1} F}^\epsilon (x) := 
            \left\lbrace
                y \in \RR^n \left | \; 
                        \mathcal M^{1/\kappa}_F (y; x) - 
                        \mathcal M^*_{F, 1/\kappa}(x) \le \epsilon
                \right.
            \right\rbrace. 
        \end{align*}
    \end{definition}
    \begin{remark}
        Setting $\epsilon = 0$, we have the exact definition of the exact resolvent given as $\mathcal J_{\beta^{-1}F}y = \mathcal J^0_{\beta^{-1}F}y$. 
    \end{remark}
    \textbf{The remainder of this section is organized as the follow}. 
    Section \ref{ssec:catalyst-intro} gives the main ideas behind the Catalyst Acceleration Framework. 
    It follows Lin et al.'s first paper \cite{lin_universal_2015} on Catalyst Acceleration and summarizes the convergence analysis of the algorithm. 
    Section \ref{ssec:catalyst-second-paper} introduces Lin et al.'s second paper \cite{lin_catalyst_2018} on Catalyst which considers the improved relative error condition stated in Definition \ref{def:catalyst-termination-c2}. 
    The relative error condition improves the total complexity and inspired nonconvex gradient based 4WD Catalyst by Paquette et al. \cite{paquette_catalyst_2018}.

@@

        Inspired by Accelerated Proximal Point Method (APPM) from Güler \cite{guler_new_1992}, and Inexact Proximal Point Method (IPPM) of Rockafellar 1976 \cite{rockafellar_monotone_1976}, Lin et al. \cite{lin_universal_2015} proposed a generic method taking inspirations from the convergence claims of APPM to accelerate Variance Reduced Method (VRM). 

@@

            \begin{theorem}[Outer loop convergence not necessarily strongly convex]\label{thm:erro-seq-outer-cnvx}\;\\

@@

                Assume that $F$ is convex but with strong convexity constant $\mu \ge 0$. 

@@

                VRMs listed earlier satisfy this assumption. 

@@

            \begin{proposition}[Inner loop convergence not necessarily strongly convex]\label{prop:inner-loop-complexity-cnvx}\;\\
